---
title: "Data Tidying"
author: "Gary Holton"
date: "1/16/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, root.dir = '~/Downloads/NCEAS/git/adc/')
```

Load packages for this session (suppress warming message)
```{r,  message=FALSE}  
library(dplyr)
library(tidyr)
```
To call a funciton from a specific package use: package_name::function_name()

#Data Cleaning

Read in data files
```{r}
catch <- read.csv('https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1',
                  stringsAsFactors = FALSE)
head(catch)

regions <- read.csv('https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1', 
                    stringsAsFactors = FALSE)
head(regions)
```

Begin data cleanup

* remove marginal sum and notes col
* move from wide to long format
```{r}
catch_long <- catch %>% 
  select(-All,-notesRegCode) %>%  
  gather(key="species", value="catch", -Year, -Region) 
    # don't gather on year or region
    # could also just specify species columns to gather on
head(catch_long)
```

Check for bad data
```{r, eval=F, echo=F}
test_catch <- as.integer(catch_long$catch)
catch_long[which(is.na(test_catch) == T),]
```

```{r}
catch_cleaned <- catch_long %>%
  rename(catch_thousands = catch) %>%
  mutate(catch_thousands = ifelse(catch_thousands == "I", 1, catch_thousands)) %>%
  mutate(catch_thousands = as.integer(catch_thousands)) %>%
  mutate(catch = catch_thousands * 1000)
```



compare sum to "all"" column

```{r, eval=F, echo=F}
catch_sum <- catch %>%
  mutate(Chinook = ifelse(Chinook == "I", 1, Chinook)) %>%
  mutate(Chinook = as.integer(Chinook) ) %>%
  mutate(sum = Chinook + Coho + Pink + Sockeye + Chum) %>%
  rename(all = All) %>%
  select(Region, Year, sum, all)

head(catch_sum)
```


Check what species exist: (this wasn't in the excercise)
```{r}
summarize(group_by(catch_cleaned,species),n())
```



#Split-Apply-Combine

Calculate total catch by region
```{r}
catch_total <- catch_cleaned %>%
  group_by(Region) %>%
  summarize(catch_region = sum(catch))
catch_total

```

The function n() with no arguments gives count within each group
```{r}
catch_total_obs <- catch_cleaned %>%
  group_by(Region) %>%
  summarize(catch_region = sum(catch), n_obs=n() )
catch_total_obs

```


Calculate yearly means
```{r}
catch_yearly <- catch_cleaned %>%
  group_by(Year) %>%
  summarize(catch_year = as.integer(mean(catch)), n())
catch_yearly
```

Filter for one species:
```{r}
catch_chinook <- catch_cleaned %>%
  filter(species == "Chinook")
head(catch_chinook)
```

Filter for one species in a particular region:
```{r}
catch_chinook_SSE <- catch_cleaned %>%
  filter(species == "Chinook" & Region == "SSE")
head(catch_chinook_SSE)
```

Change sort order in data frame:
```{r}
catch_chinook_SSE <- catch_chinook_SSE %>% arrange(-Year)
head(catch_chinook_SSE)
```



#Joins

First get the region definitions
```{r}
regions_clean <- regions %>%
  select(code, mgmtArea)
```

```{r}
catch_joined <- left_join(catch_cleaned, regions_clean, 
                          by=c("Region" = "code") )
head(catch_joined)
```


