---
title: "Data Visualization Tutorial"
author: "Gary Holton"
date: "1/17/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load packages:
```{r message=F}
library(leaflet)
library(ggplot2)
library(dplyr)
library(tidyr)
library(DT)   # dynamic tables
```

## Arctic Data Training

### Data Visualization

Data from:
https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Af119a05b-bbe7-4aea-93c6-85434dcb1c5e


```{r}
esc <- read.csv('https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Af119a05b-bbe7-4aea-93c6-85434dcb1c5e', stringsAsFactors = F)

```

Calculate annual, regional, total salmon escapements

* `separate` the date into year, month, day
* `group_by` and `summarize` to get sum over unique region-year-species combinations
* `filter` out non-salmon species
```{r}
annual <- esc %>%
  separate(sampleDate, into = c("year","month", "day")) %>%
  mutate(year = as.integer(year)) %>%
  group_by(SASAP.Region, Species, year) %>%
  summarize(escapement =sum(DailyCount)) %>%
  filter(Species %in% c("Chinook", "Sockeye", "Pink", "Chum", "Coho"))
#head(annual)
```





Make a dynamic table:
```{r}
datatable(annual)
```
## Static Plots

The R package colorbrewer can customize colors
```{r}
ggplot(data = annual, mapping = aes(x = Species, y = escapement, fill = SASAP.Region )) +
  geom_col()
```

```{r}
ggplot( filter(annual, SASAP.Region == "Kodiak"), 
    aes(year, escapement, color=Species)) +
     geom_point() + 
     xlab("Year") + ylab("Escapement") +
    ggtitle("Kodiak Salmon Escapement")
    #theme(plot.title = element_text(h))

```

### Time series

Plot time series of escapement by species for all regions
```{r, fig.height = 10, fig.width=6}
my_theme <- theme_bw() + theme(plot.title = element_text(hjust=0.5))
ggplot(annual, aes(year,escapement, color=Species)) +
  geom_line() +
  geom_point() +
  my_theme +
  facet_wrap(~SASAP.Region, scales="free_y", ncol=2) 
#ggsave("region_escapement.pdf", width=8.5, height=11, units="in")
```
Package gridExtra can be used for arranging plots of different types together (as opposed to `facet_wrap` which puts all of the same kind of plot together)





## Maps
Read in location data and use `drop_na` to get rid of rows with missing data. (Warning: this would also drop rows where Location name is NA)
```{r}
locations <- esc %>%
  distinct(Location, Latitude, Longitude) %>%
  drop_na()
```

Not create the map
```{r}
leaflet(locations) %>%
  addTiles() %>%
  addMarkers(lng=~Longitude, lat = ~Latitude, popup = ~Location)
```



